---
title: "Fundamentals of Programming 2026"
author: "Berry Boessenkool; "
date: last-modified
format: html
---

This is a markdown / github task in the course [FP26](https://open.hpi.de/courses/hpi-dh-fprog2026).  
*It is fine to not understand the R/Python code at this point - we'll get to that throughout the course :)*  

- Please first read the i.2 homework slide and follow the [installation guide](https://github.com/brry/fpsetup#software-installation-guide).
- Please **pull** before you work and again before you **push**!! We want to reduce version conflicts as much as possible.
- Include the generated md file (+ image files) in your commit.
- Write an informative commit message.
- `Pull` before you `push` your commit. Yes, this instruction is repetitive. Experience teaches it is necessary. DO NOT UPLOAD YOUR CHANGES WITHOUT FIRST UPDATING THE CODE!

## Homework

Implement one of these ideas, following the instructions above.  
To learn more markdown, create a copy of the file in which you can implement many/all ideas. Each student should have some options available, so don't push those changes.  

- change the date format in the yaml header to include HH:mm after the day (*this one should be implemented early, please*)
- change this bullet list to checkboxes for ideas that are implemented
- add some reasonable markdown code (e.g. strikethrough, link, reference, bold, cursive, ...)
- in Rstudio: change the chunk output location setting to "console" (changes yaml header)
- make one single code chunk foldable (folded by default); leave all other chunks as is
- suppress the output of `updateRdwd` with a chunk option
- improve the python output to not be split up (with a chunk option)
- add a `knitr::kable` with a small subset of the dataset
- format a plot as actual figures and add labels (chunk options)
- elsewehere in the text, add a reference to that figure
- change the date to the rendering time (note why in the commit - there is one particular reason)
- add a table of contents (yaml header option)
- set a custom theme or highlight style in the yaml header
- add a footnote somewhere in the text
- add an image (e.g. a screenshot or the DWD logo) with alt text
- set global execution options in the yaml (e.g. `warning: false`) instead of per-chunk
- add `fig-width` / `fig-height` as global defaults in the yaml
- add margin content with `column: margin` on a chunk or text block

::: {.callout-tip collapse="true"}
## click here for further ideas

These suggestions are less applicable to the document at hand or not so interesting for learning the markdown basics

- use a tabset (`:::{.panel-tabset}`) to organise sections or outputs
- add cross-references between sections using `@sec-` labels
- use a blockquote (`>`) for a fun or relevant quote about weather/data
- add an explanation why the code chunks in this particular document should not be cached
- use a meaningful multi-line commit message (subject + body) for a commit
- change some other document setting you like from [quarto.org](https://quarto.org/)
- revert a change someone made that you dislike (commit/PR discussion welcome!)
- tag a version with `git tag` after a meaningful milestone
- add a `.gitignore` entry for irrelevant output folders (e.g. `*_cache/`)
- enable `embed-resources: true` so the HTML renders as a single self-contained file
- improve a code chunk with some meaningful adaptation (this is not markdown related)
- use code annotation (`# <1>` style) to explain a line in a chunk without touching the code logic
- use a different style Quarto callout block for important info
:::

## Check package availability

```{python, modules}
import matplotlib, pandas
```

```{r, packages} 
if(!requireNamespace("rdwd", quietly=TRUE))
	install.packages("rdwd")
rdwd::updateRdwd()
```

## Get weather data

download recent weather data using [rdwd](https://brry.github.io/rdwd/)

```{r, get_weather_data}
library(rdwd)
link <- selectDWD("Potsdam", res="daily", var="kl", per="recent")
clim <- dataDWD(link, varnames=TRUE, force=24, quiet=TRUE)
```

## Visualise recent temperature

```{r, plot_clim}
plotDWD(clim, "TMK.Lufttemperatur", main=range(clim$MESS_DATUM))
```

## Transfer to Python

```{python, histograms, results="hide"}
clim_py = r.clim
print(f"Dataset shape: {clim_py.shape[0]} rows, {clim_py.shape[1]} columns")
clim_py = clim_py.select_dtypes(include=['float64', 'int64'])
clim_py.hist(figsize=(20, 16), bins=5)
```

## Summary statistics

```{python, temperature_stats}
# Calculate and display some basic statistics
print("\n=== Summary Statistics for Temperature ===")
temp_stats = clim_py['TMK.Lufttemperatur'].describe()
print(temp_stats)
```

