---
title: "Fundamentals of Programming 2026"
author: "Berry Boessenkool; "
date: now
date-format: YYYY-MM-DD, HH:mm
format: gfm
editor_options: 
  chunk_output_type: console
---

This is a github task in the course [FP26](https://open.hpi.de/courses/hpi-dh-fprog2026).  
*it is fine to not understand the code at this point - we'll get to that throughout the course :)*  

- Please follow the [installation guide](https://github.com/brry/fpsetup#software-installation-guide) first.
- Please **pull** before you work and again before you **push**!!
- Include the generated md file (+ image files) in your commit.

Ideas for the homework:

- improve a code chunk (meaningful adaptation)
- make one single code chunk foldable (folded by default). But leave all other chunks alone
- suppress the output of updateRdwd with a chunk option
- silence the `dataDWD` call (the function has an argument for that)
- improve the python output to not be split up
- revert a change someone made that you dislike (commit/PR discussion welcome!)
- add some reasonable markdown code (like strikethrough for ideas that are implemented)

## Check package availability

```{python, modules}
import matplotlib, pandas
```

```{r, packages} 
if(!requireNamespace("rdwd", quietly=TRUE))
	install.packages("rdwd")
rdwd::updateRdwd()
```

## Get weather data

download recent weather data using [rdwd](https://bookdown.org/brry/rdwd/)

```{r, get_weather_data}
library(rdwd)
link <- selectDWD("Potsdam", res="daily", var="kl", per="recent")
clim <- dataDWD(link, varnames=TRUE, force=24)
```

## Visualise recent temperature

```{r, plot_clim}
plotDWD(clim, "TMK.Lufttemperatur", main=range(clim$MESS_DATUM))
```

## Transfer to Python

```{python, histograms, results="hide"}
clim_py = r.clim
print(f"Dataset shape: {clim_py.shape[0]} rows, {clim_py.shape[1]} columns")
clim_py = clim_py.select_dtypes(include=['float64', 'int64'])
clim_py.hist(figsize=(20, 16), bins=5)
```

## Summary statistics

```{python, temperature_stats}
# Calculate and display some basic statistics
print("\n=== Summary Statistics for Temperature ===")
temp_stats = clim_py['TMK.Lufttemperatur'].describe()
print(temp_stats)
```

